\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{titlesec}

\titleformat{\section}{\large\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection.}{0.5em}{}

\title{\textbf{Kian: A Fully Offline Voice Assistant for Children's Education}}
\author{Aaron Schweiger, 2026}
\date{\vspace{-2em}}

\begin{document}
\maketitle

\begin{abstract}
Kian is an open-source, fully offline voice assistant designed for children's education.
Running entirely on an NVIDIA Jetson Orin, it requires no internet connection and no parent-managed accounts---children interact through speech alone.
A multi-stage content filtering pipeline reduces exposure to harmful material, while a heuristic LaTeX-to-speech post-processing system makes audio-based mathematical question answering feasible.
This paper describes the architecture, design rationale, and key tradeoffs of the system.
\end{abstract}

\section{Motivation}

Voice assistants have become a natural interface for children to explore questions, practice concepts, and satisfy curiosity.
However, existing solutions depend on cloud services, raising concerns about privacy, availability, and access control.

\begin{itemize}[nosep]
  \item \textbf{Privacy.} Children's voice data is sensitive. Cloud-based systems transmit raw audio to remote servers, creating risks under regulations such as COPPA. An offline system ensures that no audio or conversation data ever leaves the device.
  \item \textbf{Connectivity.} Many educational settings---rural schools, outdoor classrooms, vehicles, areas with unreliable infrastructure---lack consistent internet access. A fully on-device system operates anywhere.
  \item \textbf{Autonomy.} Children should be able to interact with an educational tool without needing a parent's phone, a managed account, or an app store. Kian is controlled entirely through speech: changing the persona, adjusting volume, setting grade level, and resetting context are all voice commands.
\end{itemize}

\section{System Architecture}

Kian is structured as an asynchronous pipeline of five stages, all running on-device:

\begin{enumerate}[nosep]
  \item \textbf{Voice Activity Detection (VAD).} Silero VAD (ONNX, ${\sim}2$\,MB) monitors the microphone stream at 48\,kHz, downsampled to 16\,kHz for inference. Speech segments are extracted with a 1.2\,s silence threshold. A 30-second maximum prevents unbounded memory growth from sustained noise.
  \item \textbf{Speech-to-Text (STT).} Faster-whisper with the \texttt{small.en} model transcribes detected speech segments using int8 quantization.
  \item \textbf{Language Model (LLM).} Qwen3-4B, quantized to Q4\_K\_M (4-bit), runs via Ollama with full GPU offload. The 2048-token context window is managed by dropping oldest conversation turns when history exceeds 1500 tokens. A smaller Qwen3.5-2B model via llama.cpp is available as a lower-latency alternative.
  \item \textbf{Content Filtering.} A streaming n-gram detector checks each token as it is generated. On match, generation halts immediately and the conversation context is reset.
  \item \textbf{Text-to-Speech (TTS).} Piper TTS synthesizes speech from text fragments, streamed to a persistent ALSA output to avoid underruns. A LaTeX-to-speech post-processor converts mathematical notation before synthesis.
\end{enumerate}

\noindent The pipeline is fully streaming: TTS begins speaking the first sentence while the LLM is still generating subsequent ones. Barge-in detection monitors microphone levels during playback, allowing the child to interrupt at any sentence boundary.

\section{Hardware Platform}

Kian targets the NVIDIA Jetson Orin Nano (8\,GB unified memory). The choice is driven by two properties of interactive voice systems:

\begin{itemize}[nosep]
  \item \textbf{Time-to-first-token (TTFT) dominates perceived latency.} In a conversational loop, the child is waiting from the moment they stop speaking until the assistant begins responding. Token generation rate (tok/s) matters less---even modest throughput produces speech faster than it can be listened to. What matters is how quickly the first token arrives.
  \item \textbf{GPU acceleration with large memory bandwidth.} The Jetson Orin's GPU and unified memory architecture provide the memory bandwidth needed to keep TTFT low for quantized transformer inference, without the overhead of discrete GPU memory transfers.
\end{itemize}

\noindent Running headless (without a desktop environment) frees 1--1.3\,GB of RAM, improving GPU layer offload. The total memory budget is approximately:

\begin{table}[h]
\centering
\begin{tabular}{lr}
\toprule
Component & Memory \\
\midrule
OS / system (headless) & $\sim$1.0\,GB \\
Ollama + LLM (Qwen3-4B Q4\_K\_M) & $\sim$3.4\,GB \\
Safety classifier (CPU) & $\sim$0.4\,GB \\
STT (Whisper tiny) & $\sim$0.1\,GB \\
VAD (Silero) + TTS (Piper) & $\sim$0.2\,GB \\
KV cache (2K context) & $\sim$0.3\,GB \\
\midrule
Headroom & $\sim$2.4\,GB \\
\bottomrule
\end{tabular}
\end{table}

\section{Speech-Only Control}

A core design principle is that no screen, keyboard, or companion device is required. All configuration is performed through natural speech commands:

\begin{itemize}[nosep]
  \item \textbf{Identity:} ``Your name is Sophie'' or ``I'll call you Max'' changes the assistant's persona and persists the choice across sessions.
  \item \textbf{Grade level:} ``I'm in second grade'' or ``I'm in grade 5'' adjusts vocabulary and explanation complexity. Supports spoken numbers (``two'', ``second'') and digits alike, for grades pre-K through 12.
  \item \textbf{Volume:} ``Whisper'', ``inside voice'', ``speak up'', and ``shout'' map to four volume presets (30\%, 50\%, 80\%, 100\%) controlled via PulseAudio.
  \item \textbf{Context:} ``Let's start fresh'' or ``forget everything'' resets conversation history. Eight minutes of silence triggers an automatic reset.
  \item \textbf{Voice:} ``Change your voice'' or ``can I talk to someone else'' switches to a different TTS voice.
\end{itemize}

\noindent All settings---name, grade, voice, and volume---are persisted to disk and restored on startup.

\section{Content Safety}

Children's content filtering requires multiple layers, since no single mechanism is sufficient:

\begin{enumerate}[nosep]
  \item \textbf{System prompt.} The LLM is instructed to behave as an educational cartoon animal speaking to a child at a specific grade level. This biases generation toward age-appropriate content.
  \item \textbf{Streaming token filter.} A sliding-window n-gram matcher checks every token against a curated list of 100+ prohibited words and phrases spanning profanity, slurs, sexual content, violence, drug references, and adult/gambling websites. Detection is immediate---the response is discarded, the context is reset, and a neutral deflection is spoken.
  \item \textbf{LLM safety classifier.} Each complete sentence is classified by a separate small language model (Qwen2.5-0.5B, Q2\_K quantization, running on CPU) before it is sent to the TTS pipeline. The classifier is prompted to flag profanity, insults, violence, weapons, sexual content, self-harm, drugs, and adult topics, responding with a single SAFE or UNSAFE token. Unsafe sentences halt generation and reset the conversation context. Running on CPU avoids contention with the primary LLM on the GPU; inference takes approximately 90\,ms per sentence. A throttle mechanism ensures safety checks are interleaved with TTS playback rather than executed in bursts.
\end{enumerate}

\noindent The n-gram filter list is maintained as a plain text file, making it straightforward for educators and parents to review and extend. The LLM classifier provides contextual understanding that rule-based filters cannot---for example, detecting instructions for harmful activities composed entirely of individually benign words.

\section{Mathematical Post-Processing}

Audio-based educational assistants face a fundamental challenge with mathematics: spoken language is inherently sequential and ambiguous for mathematical expressions. ``a plus b over c'' could mean $a + \frac{b}{c}$ or $\frac{a+b}{c}$.

Kian addresses this by allowing the LLM to emit standard LaTeX notation, which is converted to spoken English before TTS synthesis. The system:

\begin{itemize}[nosep]
  \item Detects math regions via standard delimiters (\texttt{\$...\$}, \texttt{\$\$...\$\$}, \verb|\(...\)|, \verb|\[...\]|, and \verb|\begin{equation}|).
  \item Converts fractions, roots, superscripts, subscripts, Greek letters, operators, set theory notation, calculus operators, and quantifiers.
  \item Applies disambiguation rules: compound sub-expressions (those containing addition or subtraction) are wrapped with ``the quantity \ldots end quantity'' in contexts where boundaries are ambiguous (e.g., square roots), while fractions use ``over'' as a natural delimiter.
  \item Distinguishes unary negative from binary minus: $-x$ becomes ``negative x'' while $a - b$ becomes ``a minus b''.
  \item Separates adjacent single-letter variables (e.g., $ix$ becomes ``i x'') to prevent TTS from reading them as words, while preserving known multi-letter tokens like function names.
  \item Leaves prose outside math delimiters completely untouched, so punctuation like ``Hey there!'' is never misinterpreted as factorial.
\end{itemize}

\noindent This makes it feasible for a child to ask ``what is the quadratic formula?'' and hear a correct, unambiguous spoken rendering of $x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}$.

\section{Latency Considerations}

In interactive voice applications, the relevant metric is \emph{turn latency}: the time from when the child stops speaking to when the assistant begins responding. This decomposes as:

$$t_{\text{turn}} = t_{\text{silence}} + t_{\text{STT}} + t_{\text{TTFT}} + t_{\text{TTS,first}}$$

\noindent where $t_{\text{silence}}$ is the VAD silence detection window (1.2\,s), $t_{\text{STT}}$ is transcription time, $t_{\text{TTFT}}$ is the LLM's time to first token, and $t_{\text{TTS,first}}$ is the time to synthesize and begin playing the first sentence fragment.

Crucially, \emph{tokens per second} is not the bottleneck. Even at moderate generation rates, text accumulates faster than it can be spoken. The streaming architecture exploits this: TTS begins on the first sentence boundary (minimum 25 characters) while the LLM continues generating.

The Jetson Orin's value proposition is minimizing $t_{\text{TTFT}}$: its GPU and high memory bandwidth accelerate the prefill phase of transformer inference, which is the compute-intensive step that determines when the first token appears.

\section{Wikipedia Context Augmentation}

Small language models hallucinate frequently on factual questions---a serious issue for an educational tool.
Kian mitigates this with a retrieval-augmented generation (RAG) pipeline backed by Simple English Wikipedia, an edition written at a reading level appropriate for children.

\subsection{Offline Index}

A build script (\texttt{build-wiki-db.py}) downloads the Simple Wikipedia XML dump (${\sim}350$\,MB) and constructs a local SQLite database with an FTS5 full-text search index (${\sim}170$\,MB).
The entire index lives on-device; no network access is needed at query time.

\subsection{Topic Extraction}

Before each LLM turn, the child's utterance is matched against a set of 20+ regex patterns designed to extract educational topics from natural speech: ``what is a black hole'', ``tell me about volcanoes'', ``who was Marie Curie'', ``how does photosynthesis work'', and similar phrasings.
Stopwords are stripped, and the remaining terms are used as an FTS5 query.

\subsection{Retrieval and Injection}

If the FTS5 search returns a result above a BM25 relevance threshold, the first 500 characters of the matching article are injected into the system prompt as reference material for that turn.
The LLM is instructed to use the reference only if relevant and never to mention its source.
When conversation turns are evicted from the context window, corresponding wiki references are also evicted, keeping the context clean.

This approach grounds the LLM's responses in factual content without requiring a larger model or internet access.

\section{Model Evaluation}

\subsection{Methodology}

We evaluated candidate models using a sequential prompt benchmark that simulates a child's conversation.
Eight prompts progress from a simple greeting through factual questions to a creative storytelling request, building conversational context incrementally.
Each model was assessed on three qualitative dimensions:

\begin{itemize}[nosep]
  \item \textbf{Succinctness}: Are short-answer responses concise? (Evaluated on 6 of the prompts.)
  \item \textbf{Complexity}: Is the storytelling response sufficiently detailed? (Evaluated on 1 creative prompt.)
  \item \textbf{Accuracy}: Are responses factually correct and free of hallucinations? (Evaluated across all 8 prompts.)
\end{itemize}

\noindent Models were benchmarked on the target hardware (Jetson Orin Nano, 8\,GB) using two runtimes:
llama-cpp-python\footnote{\url{https://github.com/abetlen/llama-cpp-python}} for in-process GGUF inference, and
Ollama\footnote{\url{https://ollama.com/}} as an OpenAI-compatible server with native CUDA support.

\subsection{Performance Results}

All models were evaluated with a 2048-token context window across 10 benchmark runs (80 prompts per model).
Repeating the same prompt sequence multiple times averages over two sources of variance: randomness in the model's own generated responses (due to sampling temperature), and randomness in the LLM-as-judge evaluation of qualitative scores.
Table~\ref{tab:perf} summarizes latency and throughput; Table~\ref{tab:quality} summarizes qualitative response quality.

\begin{table}[h]
\centering
\small
\input{benchmark-table}
\caption{Performance benchmarks on Jetson Orin Nano (8\,GB, headless). Mean and p95 TTFT over 10 runs $\times$ 8 prompts = 80 samples per engine. All models use Q4\_K\_M quantization and 2048-token context. RAM is total allocation reported by the runtime including KV cache.}
\label{tab:perf}
\end{table}

\begin{table}[h]
\centering
\small
\input{quality-table}
\caption{Qualitative response quality across 10 benchmark runs (8 prompts each). All metrics are lower-is-better: Err = factual errors per conversation, Chatty = verbose or repetitive responses (of 5), Poor Prose = weak explanatory/story responses (of 3).}
\label{tab:quality}
\end{table}

\subsection{Model Sources}

\begin{itemize}[nosep]
  \item Qwen3.5-2B Q4\_K\_M GGUF: \url{https://huggingface.co/unsloth/Qwen3.5-2B-GGUF}
  \item Qwen3-4B Q4\_K\_M: \url{https://ollama.com/library/qwen3}
  \item Qwen3.5-2B, Qwen3.5-4B: \url{https://ollama.com/library/qwen3.5}
  \item Llama 3.2-3B Instruct Q4\_K\_M: \url{https://ollama.com/library/llama3.2}
  \item Ministral-3 3B: \url{https://ollama.com/library/ministral-3}
  \item Granite 3.3-2B: \url{https://huggingface.co/ibm-granite/granite-3.3-2b-instruct-GGUF}
  \item Granite 4.0 Micro IQ4\_XS: \url{https://huggingface.co/bartowski/ibm-granite_granite-4.0-micro-GGUF}
  \item Granite 4-3B: \url{https://ollama.com/library/granite4}
  \item Nemotron-3 Nano 4B: \url{https://ollama.com/library/nemotron-3-nano}
  \item Silero VAD: \url{https://github.com/snakers4/silero-vad}
  \item Faster-whisper: \url{https://github.com/SYSTRAN/faster-whisper}
  \item Piper TTS: \url{https://github.com/rhasspy/piper}
  \item Simple English Wikipedia: \url{https://dumps.wikimedia.org/simplewiki/}
\end{itemize}

\section{Limitations and Future Work}

\begin{itemize}[nosep]
  \item \textbf{Language support} is currently English-only across all pipeline stages.
  \item \textbf{Content filtering} is rule-based and cannot catch novel circumventions or contextually inappropriate content that uses benign words. Adversarial probing by children is a realistic threat model.
  \item \textbf{Mathematical rendering} handles standard LaTeX but not all edge cases (e.g., matrices, piecewise functions, or deeply nested expressions).
  \item \textbf{No visual output} means that some mathematical concepts (geometry, graphs) are difficult to convey through speech alone.
\end{itemize}

\section{Conclusion}

Kian demonstrates that a practical, fully offline educational voice assistant can run on modest edge hardware. By keeping all processing on-device, it provides strong privacy guarantees and operates without internet access. Speech-only control eliminates the need for screens or parent-managed accounts. Multi-layered content filtering and LaTeX-to-speech post-processing address two key challenges of building voice interfaces for children: safety and mathematical communication.

\end{document}
