\begin{table}[h]
\centering
\small
\begin{tabular}{l|r|r|r|r}
\toprule
Engine & Overall & 4th-Wall & Fact Err Avg & Fact Err Max \\
\midrule
ollama:Ministral-3 3B & 3.2 & 0.0 & 3.8 & 5 \\
server:qwen3-4b-instruct-2507-q4\_k\_m & 3.0 & 0.6 & 3.0 & 4 \\
ollama:gemma3:4b & 3.0 & 0.0 & 3.4 & 5 \\
server:granite-4.0-h-micro-Q4\_K\_M-c1024 & 2.8 & 1.4 & 3.0 & 4 \\
ollama:Qwen3-4B & 2.8 & 1.0 & 3.6 & 5 \\
server:granite-4.0-h-micro-Q4\_K\_M & 2.6 & 1.0 & 2.6 & 3 \\
ollama:Llama 3.2-3B & 2.4 & 0.4 & 3.8 & 6 \\
server:ibm-granite\_granite-4.0-micro-IQ4\_XS & 2.2 & 1.0 & 2.6 & 4 \\
ollama:Granite 4-3B & 2.2 & 1.0 & 3.0 & 4 \\
server:granite-4.0-micro-Q4\_K\_M & 2.2 & 0.8 & 3.0 & 5 \\
server:Bonsai-8B & 2.2 & 1.0 & 4.0 & 5 \\
ollama:Granite 3.3-2B & 2.0 & 0.8 & 4.2 & 7 \\
server:granite-3.3-2b-instruct-Q4\_K\_M & 2.0 & 1.0 & 4.2 & 5 \\
ollama:Nemotron-3 Nano 4B & 1.8 & 0.8 & 5.2 & 6 \\
server:Qwen3.5-2B-Q4\_K\_M & 1.8 & 0.2 & 6.6 & 10 \\
ollama:Qwen3.5-2B & 1.0 & 0.6 & 6.8 & 8 \\
\bottomrule
\end{tabular}
\caption{Qualitative response quality across 5 benchmark runs (20 prompts each; two single-token cache-warming probes are excluded). Overall is a 1--5 holistic score (higher is better); 4th-Wall counts responses breaking the cartoon-animal persona; Fact Err counts factual errors on factual prompts (lower is better). Evaluated by Claude Opus 4.6.}
\label{tab:quality}
\end{table}

\begin{table}[h]
\centering
\small
\begin{tabular}{l|l}
\toprule
Engine & Representative flub \\
\midrule
ollama:Ministral-3 3B & \textit{``blue blood (which makes them glow in the dark!)''} \\
server:qwen3-4b-instruct-2507-q4\_k\_m & \textit{``a duck, not a bird''} \\
ollama:gemma3:4b & \textit{``they have a total of seven arms''} \\
server:granite-4.0-h-micro-Q4\_K\_M-c1024 & \textit{``I'm a friendly cartoon animal''} \\
ollama:Qwen3-4B & \textit{``same big family of animals that include dolphins, whales, and even sharks''} \\
server:granite-4.0-h-micro-Q4\_K\_M & \textit{``like elephants or elephants''} \\
ollama:Llama 3.2-3B & \textit{``The fastest land animal isn't even an animal''} \\
server:ibm-granite\_granite-4.0-micro-IQ4\_XS & \textit{``became fish, like dolphins, sharks, and whales''} \\
ollama:Granite 4-3B & \textit{``I'm Kian, your friendly cartoon animal helper''} \\
server:granite-4.0-micro-Q4\_K\_M & \textit{``a special ring that they use to throw things''} \\
server:Bonsai-8B & \textit{``God placed a pear-shaped object in Adam's throat''} \\
ollama:Granite 3.3-2B & \textit{``are fully terrestrial animals, living in oceans''} \\
server:granite-3.3-2b-instruct-Q4\_K\_M & \textit{``This area is actually the grasshopper muscle''} \\
ollama:Nemotron-3 Nano 4B & \textit{``eight hearts pump blood to each arm''} \\
server:Qwen3.5-2B-Q4\_K\_M & \textit{``Orcas are definitely not whales because they do not have a backbone''} \\
ollama:Qwen3.5-2B & \textit{``a human is not a mammal or even an animal at all''} \\
\bottomrule
\end{tabular}
\caption{Representative factual errors from qualitative evaluation.}
\label{tab:flubs}
\end{table}

\textbf{Ministral-3 3B} achieved the highest overall score (3.2) while maintaining a perfect fourth-wall record, never once breaking character across five conversations. \textbf{Gemma3 4B} matched this persona discipline but trailed slightly in overall quality (3.0). \textbf{Qwen3-4B-Instruct-2507} (served via llama.cpp) rounded out the top tier at 3.0 overall with the joint-lowest average factual error rate among high-scoring models (3.0), though it occasionally broke character. An interesting tension appears with the \textbf{Granite 4.0 Tiny} variants: the llama.cpp serve achieved one of the lowest factual error rates in the field (2.6 avg, 3 max) but scored only 2.6 overall, suggesting that factual caution alone does not guarantee engaging conversation. The reduced-context variant (\textbf{Granite 4.0 Tiny} at $c{=}1024$, half the default 2048-token window) scored slightly higher overall (2.8 vs 2.6) but posted the highest fourth-wall break rate in the table (1.4); the reduced window was tested as a speed and trim-cost experiment, not as an expected quality improvement. At the bottom of the ranking, \textbf{Qwen3.5-2B} was the weakest performer on both axes: the Ollama variant scored just 1.0 overall with 6.8 average factual errors, while its llama.cpp counterpart improved modestly to 1.8 overall but posted the single worst factual-error maximum in the study (10 errors in one conversation), including the confident claim that ``orcas are definitely not whales because they do not have a backbone.'' \textbf{Nemotron-3 Nano 4B} fared only marginally better (1.8), with the highest sustained factual error rate among 4B-class models (5.2 avg).

Where both Ollama and llama.cpp variants of the same base model were tested, differences were generally small. \textbf{Granite 3.3 2B} scored identically (2.0 overall, 4.2 fact avg) across both serving backends. The Qwen3.5-2B pair showed a larger gap (1.0 vs.\ 1.8), likely attributable to different quantization schemes rather than the runtime itself. For this use case---sustaining a factually grounded, in-character multi-turn conversation on a resource-constrained device---Ministral-3 3B and Gemma3 4B emerge as the strongest candidates, offering the best balance of persona consistency, conversational quality, and acceptable (though imperfect) factual accuracy.

\subsection{Benchmark Limitations}

This benchmark is a task-specific internal evaluation, not a general capability ranking. The prompt set is small (20 prompts across 5 conversations) and English-only, covering a narrow slice of the topics a child might raise. Qualitative scores were assigned by a single LLM judge (Claude Opus 4.6) rather than human raters, which may not fully capture dimensions of child-appropriateness such as tone, pacing, emotional register, or age-calibrated vocabulary. The benchmark does not evaluate the content-safety pipeline, the text-to-speech system, or latency under real serving conditions. Ollama and llama.cpp variants may use different quantization levels or prompt templates, so backend comparisons reflect the full serving configuration rather than a controlled ablation. Results should be interpreted as a comparative signal for model selection on this hardware, not as absolute quality claims.
